### 2011-05-27 thesis/contents.tex
More thesis structure
Added content from the tech report. Its not at all the final content but it helps structuring.

\autoref{chap:background} introduces background concepts for understanding the
presented thesis. In particular \autoref{sec:graphical-models} introduces
\emph{probabilistic graphical models} that lay the base modelling tool for
the semantic structured representation over which novelty detection
methods are developed.

@@

\section{Classification}
\subsection{Recognition and Categorization}
\subsection{Support Vector Machines}
\subsection{Multi-Class Classification}
%\subsection{Kernel-Trick}

\section{Visual Features}
\subsection{Local Features}
\subsection{Global Features}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Important sections of this chapter begin here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Probability Theory}

@@

% In case we expand in some finding structure in graphs maybe the following will
% be useful
\section{Label Propagation}
\section{Min-Cuts}

>>>>

\autoref{chap:background} introduces background concepts for understanding the
presented thesis. In particular \autoref{sec:graphical-models} introduces
\emph{probabilistic graphical models} that lay the base modelling tool for
the used structured semantic representation.

@@

\section{Classification}
\subsection{Recognition and Categorization}
\subsection{Support Vector Machines}
\Glspl{SVM} where introduced by \cite{cortes1995support} and can be seen as
discriminative linear based classifier. They are based on a strong mathematical
foundation and have powerful generalization capabilities. In their original form
\gls{SVM} separates two classes of points in an hyper-space with a
\emph{maximal margin hyperplane}~(\autoref{fig:svm-sample}).

Later they were extended to deal with noisy data by using \emph{soft margins}.
And to handle non-linear spaces as seen on \autoref{sec:kernel-trick}.
Although they only allow for two class-classification several methods have been
proposed to build multi-class classifiers based on binary classifiers as seen on
\autoref{sec:multiclass-classifiers}.

\begin{figure}[h]
\begin{center}
% TODO: get a decent picture to illustrate SVMs
\includegraphics[width=0.4\textwidth]{figures/Svm_max_sep_hyperplane_with_margin}
\end{center}
\caption{{SVM} separating two class of points by a
         \emph{maximal margin hyperplane}. The hyperplane can be described by
         the collection of support vectors and associated weights, marked in the
         image as sample points with large borders.}
\label{fig:svm-sample}
\end{figure}

They have been used in several classification and recognition problems and are
in fact a standard across machine learning techniques. Their efficiency, exact
training results and generalization made their suitable for many tasks. Such as
text categorization, digit-recognition, spam-classification.
They have also been extensively used in visual place classification.

\subsection{Multi-Class Classification}
\label{sec:multiclass-classifiers}
\Glspl{SVM} were designed as two-class classifier. And so they need to be
adapted for using in multi-class problems. The most common usage is to design
multi-class classifiers recurring at the usage of multiple two-class classifiers
and methods for combining them. The most typical approaches for a multi-class
classification on $c$ classes using \glspl{SVM} are:

\begin{description}
\item[One Against All] - in this method $c$ distinct classifiers are trained to
distinguish any class of the remaining ones. The output of all those classifiers
(distance to the separating hyperplane) is then used to categorize the output.
The most common approach is to pick the class with the largest hyperplane
distance. Other variations exists as is the case of using the minimal distance
to the average classification distance of each class~\citep{pronobis2007confidence}.

\item[One Against One] - in this method $c*(c-1)/2$ classifiers are trained to
distinguish between each pair of classes. The final decision is based on the
output of all those classifiers being common to use a majority vote strategy.
\end{description}

\subsection{Kernel-Trick}
\label{sec:kernel-trick}
\gls{SVM} in their basic form are only able to handle linear spaces.
Nonetheless the classes are most of the time not linearly separable in the input
space. Although there might exists a transformation $\phi$ from the original
space into a space $H$ where the input becomes linearly separable.

The Kernel trick allows to extend the \gls{SVM} definition to work on such space
$H$ without ever performing an implicit transformation between spaces. Being
enough for that to have a Kernel function defining an inner-product inside such
space: $K(x_i, x_j) = \phi(x_i)\cdot\phi(x_j)$.
In that sense a Kernel function can be seen as a function that calculates some
similarity measure between two inputs.

Several kernel functions have been proposed being the most commonly used:

\begin{description}
\item[Polynomial Kernel] - $K(x, y) = (x \cdot y + p)^d$
\item[Radial Basis Function] - $K(x, y) = e^{-\gamma\|x - y \|^2}$
\item[Histogram intersection] - $K(x, y) = e^{-\gamma \chi^2(x,y)}$, where
$\chi^2(x,y) = \sum_{i=1}^{N}\frac{(x_i-y_i)}{x_i+y_i}$ introduced by
\cite{barla2003histogram} allows to compute histogram similarity.
\item[Matching Kernels] - mimic matching similarity and are used when each
sample is represented as a set of features~\citep{boughorbel2005intermediate}.
\end{description}

\section{Features}
A feature is a piece of information which is expected to reveal information for
solving a specific task. In that sense features are task-dependant and they will
yield different performance based on the type of task they are applied to.

A wanted property on features is its repeatability under similar conditions for
the problem in hand. This is: they should be stable and invariant across
unwanted types of transformations and noise. For example a visual feature for
object detection should be present even if the target object was translated,
scaled, rotated, the light-conditions have changed or even if the object is
partially occluded.

Extracting features with those properties allows to greatly reduce the size of
input by removing unwanted noise and useless information from the captured data.
Turning the classification problem easier, more reliable and more efficient.

Often several and different types of features need to be extracted. It has been
reported by \cite{pronobis2010ijrr} that using multiple features provides a
great benefit in the context of place classification.
And \cite{quattoni2009recognizing} has showed that different types have
different impact in indoor scene recognition based on the type of scene
matching. Namely was seen that some room-categories are more likely to be
recognized by the presence of some objects and others by it generic appearance.

In the context of robotics, sensors such as cameras, laser scans are used to
sense the surrounding environment, and features can be extracted from all those.
Nonetheless as described in \autoref{sec:visual-motivation} the visual sensor is
incredibly rich and most of our features will come from it.

Visual features can be seen as belonging to two categories:
\emph{local features} and \emph{global features}.
\emph{Local features} describe fine grain properties of a part of image.
For example the existence of specific corner or an edge. \Gls{SIFT}
(\autoref{sec:sift}) is an example of such feature and has been proven useful
for matching points between images and subsequence extension to object detection.

\emph{Global features} such as \gls{CRFH} (\autoref{sec:crfh}) or
{gist}~(\autoref{sec:gist}) try to describe the whole image. Either by
statistical analysis of features over all the image or by a structured
distribution of textures findable in the image.

\subsection{Local Features}
\label{sec:local-features}

\subsubsection{SIFT}
\label{sec:sift}
The detection of interesting points has been studied for several years and is
the base of several computer vision problems solution. It allows to perform
point matching which can be used in several areas from image stitching,
3D reconstruction, video tracking, object detection, etc\dots

The most used method was presented by \cite{lowe1999object}. And its based on
building a feature vector for each image. Each of those features is based on
\emph{interesting points} detected by detecting maxima and minima of a
difference of Gaussian functions applied in a scale-space.
The scale space is used to provide scale invariant detection. Gaussian functions
are used as they are the only way to model a linear scale-space.
Each interesting point is then described by a container that is rotation
invariant.

By seeing an object as a set of features points, index and matching is then
performed by a high-dimensional search on a database of know objects. After
matching objects can be verified for geometric coherence between features.
\Gls{SVM} classifiers can also be trained to detect objects based on this type
of local features by using \emph{Matching kernels} (\autoref{sec:kernel-trick}).

\begin{figure}[h]
    \includegraphics[width=\textwidth]{figures/sift/sift.pdf}
    \caption{{SIFT} and other local features have been proven useful in object
             detection.}
\end{figure}

\subsection{Global Features}

\subsubsection{Gist of a Scene}
\label{sec:gist}
\cite{oliva2006building} argue that fast scene recognition does not need to be
built on top of object recognition but can be analyzed by scene-centered
mechanisms.
They defend that position by pointing out behaviours on human vision:
when provided with a glance of a shot a person can identify the meaning of that
given shot or "gist of a scene" without remembering specific details.

\begin{figure}[h]
\center
\includegraphics[width=0.60\textwidth]{figures/gist.jpg}
\caption{An illustration of the gist of an image. Top row: original image I;
         bottom row: noise image J for which gist(I) = gist(J). We see that the
         gist captures the dominant textural features of the overall image, and
         their coarse spatial layout~\citep{murphy2006object}.}
\end{figure}

\subsubsection{{CRFH} - Composed Receptive Field Histograms}
\label{sec:crfh}\label{sec:global-features}
Composed Receptive Field Histograms are a multidimensional statistical
representation of the occurrence of several image descriptors applied to an
image. They can be seen as an high-dimension histogram where each cell records
how many pixels of the image have the cell response for the applied descriptors.
Such high-dimensional histogram is expected to be able to better global
information contained in the image by capturing several properties of the image
as well the relations between them.

\begin{figure}[h]
\begin{center}
\includegraphics[width=1\textwidth]{figures/crfh_model.jpg}
\end{center}
\caption{A two dimensional histogram of the image built out from two image
         descriptors: $Lx$ and $Ly$. First-order Gaussian derivatives of image
         luminance in horizontal and vertical direction applied at a scale 4.}
\end{figure}

Several type of image descriptors can be applied.
For example Gaussian derivatives (such as $L_x$, $L_y$, $L_{xx}$, $L_{yy}$) are
partial derivatives of the image luminance after applying a Gaussian filter of a
given scale ($\sigma$) on the image. Gradient magnitude descriptors
(such as $|\nabla L|$) can also be used as a rotation invariant descriptor.

One of the problems of high-dimensional histograms is the memory and
computationally complexity needed to handle them.
\cite{linde2004object} suggested using a sparse form to represent those by
storing only the non-zero-cells in an array structure.

Multidimensional histograms have proven to be useful in the context of object
recognition~\citep{schiele1996object}. And have also been previously used in
the context of visual place classification~\cite{pronobis2010ijrr}.
\emph{Histogram intersection kernels} (\autoref{sec:kernel-trick}) can be used
to train classifiers based on this feature.

\subsection{Laser Features}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Important sections of this chapter begin here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Probability Theory}
\subsection{Bayes Theorem}

@@

% In case we expand in some finding structure in graphs maybe the following will
% be useful
\section{Label Propagation, Min-Cuts}
